Relational programming in Smalltalk

### 2019-03-18 booklet/article.tex
Restore some paragraphs deleted by the merge, my fault.

Reifier>>#reifyCollection: aCollection 
    ^ aCollection inject: state into: [:s :c | s reifier reify: c]
\end{minted}
requiring the auxiliary message
\begin{minted}[fontsize=\footnotesize]{smalltalk}
State>>#reifyVar: aVar 
    substitution
        at: aVar
        ifAbsentPut: [ReifiedVar new 
                        id: substitution size; 
                        yourself]
\end{minted}
and
\begin{minted}[fontsize=\footnotesize]{smalltalk}
State>>#walk: anObj 
    | k |
    k := anObj.
    [k := substitution at: k ifAbsent: [^ k]] repeat
\end{minted}

\hrulefill

where a \Verb|Walker| object is built by
\begin{minted}[fontsize=\footnotesize]{smalltalk}
Walker class>>#on: aState
    ^ self new
        state: aState;
        yourself
\end{minted}
and it responds to the selector \Verb|#walkStar| with
\begin{minted}[fontsize=\footnotesize]{smalltalk}
Walker>>#walkStar: aVar
    ^ (state walk: aVar)
        walkStarUsingWalker: self
\end{minted}
which relies on
\begin{minted}[fontsize=\footnotesize]{smalltalk}
State>>#walk: anObj
    | k |
    k := anObj.
    [k := substitution
            at: k
            ifAbsent: [^ k]] repeat
\end{minted}
and objects of different classes responds to \Verb|#walkStarUsingWalker:| as follows
\begin{minted}[fontsize=\footnotesize]{smalltalk}
Object>>#walkStarUsingWalker: aWalker
    ^ aWalker walkStarObject: self

>>>>

Reifier>>#reifyCollection: aCollection 
    ^ aCollection inject: state into: [:s :c | s reifier reify: c]
\end{minted}
requiring the auxiliary message
\begin{minted}[fontsize=\footnotesize]{smalltalk}
State>>#reifyVar: aVar 
    substitution
        at: aVar
        ifAbsentPut: [ReifiedVar new 
                        id: substitution size; 
                        yourself]
\end{minted}
and
\begin{minted}[fontsize=\footnotesize]{smalltalk}
State>>#walk: anObj 
    | k |
    k := anObj.
    [k := substitution at: k ifAbsent: [^ k]] repeat
\end{minted}
Observe that \ct{Chain} objects are also \textit{foldable} according to the
Haskell \ct{Foldable} typeclass because they respond to the message
\begin{minted}[fontsize=\footnotesize]{smalltalk}
Cons>>#inject: init into: aBlock 
    | accumulated |
    accumulated := aBlock value: init value: car.
    ^ aBlock value: accumulated value: cdr
\end{minted}

The reification process requires \ct{Walker} objects which are built by
\ct{State} objects using the template method
\begin{minted}[fontsize=\footnotesize]{smalltalk}
State>>#walker
    ^ Walker on: self
\end{minted}
and it responds to the selector \Verb|#walkStar| with
\begin{minted}[fontsize=\footnotesize]{smalltalk}
Walker>>#walkStar: aVar
    ^ (state walk: aVar)
        walkStarUsingWalker: self
\end{minted}
and objects of different classes responds to \Verb|#walkStarUsingWalker:| as follows
\begin{minted}[fontsize=\footnotesize]{smalltalk}
Object>>#walkStarUsingWalker: aWalker
    ^ aWalker walkStarObject: self
